---
title: "Attrition Case Study"
author: "Scott Lovelady"
date: "11/21/2021"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```


## Load data and packages


```{r}
library(tidyverse)
library(caret)

bi <- read.csv("bioimplants.csv")

summary(bi)

glimpse(bi)

```


### Attrition Rate


```{r}
#attrition rate
mean(bi$attrition == "Yes")

#summary table: attrition by department and job role
bi %>%
  group_by(department, job_role) %>%
  summarize(attrition = mean(attrition == "Yes")) %>%
  arrange(desc(attrition))


```

### Logistic Regression Model


```{r}
# Fit glm model using caret (to get cross-validation estimates of model's 
# out of sample performance)

set.seed(123)

#use caret for glm model
caret_mod <- train(attrition ~ ., 
                   method = "glm", 
                   data = subset(bi, select = - employee_number),
                   preProcess = c("center", "scale"))
summary(caret_mod) %>% coefficients %>% round(2)

#estimate in-sample performance
(predict(caret_mod, newdata = bi) == bi$attrition) %>% 
  mean

#out-of-sample accuracy
caret_mod

#improvement over predicting with majority class
0.871999 - mean(bi$attrition == "No")

```

> The glm model offers an improvement of 3 percentage points over predicting with the majority class.

### Leading Predictor 

```{r}
# Refit the model
set.seed(123)
caret_mod <- train(attrition ~ ., 
                   method = "glm", 
                   data = subset(bi, select = - c(employee_number, department)),
                   preProcess = c("center", "scale"))
summary(caret_mod) %>% coefficients %>% round(2)

```

> The over_time predictor has the largest effect size. The over_time coefficient of 0.89 represents an increase of 0.89 in the log odds of attrition when an employee works more than 40 hours per week.

### Policy Change

```{r}
#churn if no one works overtime
newp <- predict(caret_mod, 
        newdata = (bi %>% mutate(over_time = "No") %>% subset(select = - employee_number)),
        type = "prob")$Yes %>% 
  mean

#existing churn
existingp <- mean(bi$attrition == "Yes")

#change in percentage points if no overtime
(existingp - newp) * 100

#percent change in attrition if no overtime
(existingp - newp)  / (existingp ) * 100

#attrition by overtime
bi %>%
  group_by(over_time) %>%
  summarize(attrition = mean(attrition == "Yes")) %>%
  arrange(desc(attrition))

```

> The company should restrict employees from working more than 40 hours per week to reduce attrition. This policy change would decrease the likelihood of churn by almost 6 percentage points (a 36% decrease).

### Recommendation

> The company should restrict employees from working more than 40 hours per week. Employees that work overtime were 3 times more likely to quit in 2016 compared to employees that did not work overtime (30% attrition versus 10% attrition, respectively). Eliminating overtime would reduce the company-wide attrition rate by 36% (16% attrition with overtime versus 10% attrition without overtime).